Research Cases

### 2025-10-13 Code/R/Research/cases_research.Rmd
Start mv cases research

@@

```{r}
construct_periodics <- function(N,
                                ampl,
                                rate,
                                freq,
                                phase,
                                complex = TRUE,
                                from = 0,
                                by = 1) {
  if (any(c(length(ampl), length(rate), length(freq)) != length(phase)))
    stop("Amplitudes, rates, frequencies and phases all must have the same length")

  if (complex) {
    period_f <- exp
    unit <- 1i
  }
  else {
    period_f <- cos
    unit <- 1
  }

  colSums(ampl * exp(rate) * period_f(phase + 2 * pi * unit * freq %o% seq(
    from = from,
    length.out = N,
    by = by
  )))
}

calc_rank <- function(ampl, rate, freq, phase, complex = TRUE, tol = 1e-6) {
  if (any(c(length(ampl), length(rate), length(freq)) != length(phase)))
    stop("Amplitudes, rates, frequencies and phases all must have the same length")

  (2 - complex) *
    (sum(A != 0) + nrow(unique(matrix(c(freqs.true, rate), ncol = 2))) - length(A)) - sum(abs(freqs.true - 0.5) < tol)
}
```

# Two periodics case
## Short series
```{r}
# no rates
rates.true <- c(0, 0)
rates_case <- "_nr"
# small equal rated
# rates.true <- c(-0.01, -0.01)
# rates_case <- "_ser"
# large equal rates
# rates.true <- c(-0.02, -0.02)
# rates_case <- "_ler"
# different rated
# rates.true <- c(-0.01, -0.02)
# rates_case <- "_dr"

@@

@@

# N <- 25
# n_case <- ""
N <- 100
n_case <- "_long"

full_case <- paste0(n_case, rates_case, freqs_case, A_case, phases_case, complex_case)
max.r <- calc_rank(A, rates.true, freqs.true, phases.true, complex.signal)

stopifnot(N > max.r * 2 + 1)
signal <- construct_periodics(N, A, rates.true, freqs.true, phases.true, complex.signal)

groups_ssa <- list(1:max.r)

@@

params_grid_ssa <- list(sd = sds, L_ssa = Ls_ssa)
params_grid_hossa <- list(sd = sds,
                          td_hossa =  trunc_dims,
                          L_hossa = tens.dims_list)
```

>>>>

pred_mssa_rec <- function() {
  noise <- rnorm(N * P, 0, sd)
  x <- signal + noise
  rec <- reconstruct(ssa(x, L = L_mssa, kind = 'mssa'), groups = groups_mssa)
  Reduce("+", rec)
}

pred_homssa_rec <- function() {
  noise <- rnorm(N * P, 0, sd)
  x <- signal + noise

  rec <- tens_mssa_reconstruct(
    x,
    L_homssa,
    groups = groups_homssa,
    groups3 = groups3_homssa,
    decomp = "hooi",
    status = FALSE
  )

@@

```{r}
# If parameters are matrices, then a multivariate series will be constructed.
# In this parameters matrices columns correspond to series and rows to summands
# This means, if you want to increase number of ts in a sum, you grow rows,
# and if you want to add a 1d ts to a mv ts, you grow columns.
# The results is either a 1d ts, or a matrix, where each column is a 1d ts.
construct_periodics <- function(N,
                                ampl,
                                rate,
                                freq,
                                phase,
                                complex = TRUE,
                                from = 0,
                                by = 1) {
  if (any(c(length(ampl), length(rate), length(freq)) != length(phase))) {
    stop("Amplitudes, rates, frequencies and phases all must have the same length")
  }
  if (all(is.matrix(ampl), is.matrix(rate), is.matrix(freq), is.matrix(phase))) {
    if (any(c(dim(ampl), dim(rate), dim(freq)) != dim(phase))) {
      stop("Amplitudes, rates, frequencies and phases all must have the same dimensions")
    }
  } else if (any(is.matrix(ampl), is.matrix(rate), is.matrix(freq), is.matrix(phase))) {
    stop("All parameters should be either matrices, or vectors at the same time")
  }

  if (complex) {
    period_f <- exp
    unit <- 1i
  } else {
    period_f <- cos
    unit <- 1
  }
  N_ones <- rep(1, N)
  ts_arr <- (N_ones %o% ampl) * (N_ones %o% exp(rate)) *
    period_f(N_ones %o% phase + 2 * pi * unit * seq(
      from = from,
      length.out = N,
      by = by
    ) %o% freq)

  apply(ts_arr, 1:(length(dim(ts_arr)) - 1), sum)
}

calc_rank <- function(ampl, rate, freq, phase, complex = TRUE, tol = 1e-6) {
  if (any(c(length(ampl), length(rate), length(freq)) != length(phase)))
    stop("Amplitudes, rates, frequencies and phases all must have the same length")
  fr_unique <- unique(matrix(c(freq, rate), ncol = 2))
  (2 - complex) *
    (sum(ampl != 0) + nrow(fr_unique) - length(ampl)) - sum(abs(fr_unique[,1] - 0.5) < tol)
}

# Assumes all amplitudes are non-zero
# TODO: This is totally wrong
calc_rank3 <- function(rate, freq, phase, complex = TRUE, tol = 1e-6) {
  if (any(c(length(ampl), length(rate), length(freq)) != length(phase)))
    stop("Amplitudes, rates, frequencies and phases all must have the same length")
  f_unique <- nrow(unique(freq))
  r_unique <- nrow(unique(rate))
  p_unique <- nrow(unique(phase))
  return(min(ncol(rate), f_unique + r_unique + p_unique - 2))
}
```

# Two periodics case
## Parameters for 1d series
```{r}
# no rates
rates.true <- c(0, 0)
rates_case <- "_nr"
# small equal rated
# rates.true <- c(-0.01, -0.01)
# rates_case <- "_ser"
# large equal rates
# rates.true <- c(-0.02, -0.02)
# rates_case <- "_ler"
# different rates
# rates.true <- c(-0.01, -0.02)
# rates_case <- "_dr"

@@

# complex.signal <- TRUE
# complex_case <- "_c" # complex
complex.signal <- FALSE
complex_case <- "_r" # real

# N <- 25
# n_case <- ""
N <- 100
n_case <- "_long"
```

## Parameters for mv series (P = 2)
```{r}
P <- 2
P_case <- "2s"

# no rates
rates.true <- matrix(c(0, 0, 0, 0), ncol = P)
rates_case <- "_nr"
# small equal rated
# rates.true <- matrix(c(-0.01, -0.01, -0.01, -0.01), ncol = P)
# rates_case <- "_ser"
# large equal rates
# rates.true <- matrix(c(-0.02, -0.02, -0.02, -0.02), ncol = P)
# rates_case <- "_ler"
# different by summands rates
# rates.true <- matrix(c(-0.01, -0.01, -0.02, -0.02), ncol = P)
# rates_case <- "_dsr"
# different by time series rates
# rates.true <- matrix(c(-0.01, -0.02, -0.01, -0.02), ncol = P)
# rates_case <- "_dtr"
# different rates
# rates.true <- matrix(c(-0.01, -0.02, -0.015, -0.025), ncol = P)
# rates_case <- "_dr"

# no freqs
# freqs.true <- matrix(c(0, 0, 0, 0), ncol = P)
# freqs_case <- "_nf"
# different by summands freqs
freqs.true <- matrix(c(0.2, 0.2, 0.22, 0.22), ncol = P)
freqs_case <- "_dsf"
# different by time series freqs
# freqs.true <- matrix(c(0.2, 0.22, 0.2, 0.22), ncol = P)
# freqs_case <- "_dtf"
# different freqs
# freqs.true <- matrix(c(0.2, 0.22, 0.18, 0.24), ncol = P)
# freqs_case <- "_df"
# close different by summands freqs
# freqs.true <- matrix(c(0.02, 0.02, 0.0205, 0.0205), ncol = P)
# freqs_case <- "_cdsf"

# equal amplitudes
A <- matrix(c(1, 1, 1, 1), ncol = P)
A_case <- "_ea"
# different amplitudes
# A <- matrix(c(1, 2, 1.5, 2.5), ncol = P)
# A_case <- "_da"

# equal phases
phases.true <- matrix(c(0, 0, 0, 0), ncol = P)
phases_case <- "_ep"
# different by time series phases
# phases.true <- matrix(c(0, pi / 3, 0, pi / 3), ncol = P)
# phases_case <- "_dtp"

@@

N <- 25
n_case <- ""
# N <- 100
# n_case <- "_long"
```

```{r}
full_case <- paste0(n_case, rates_case, freqs_case, A_case, phases_case, complex_case)
max.r <- calc_rank(A, rates.true, freqs.true, phases.true, complex.signal)

stopifnot(N > max.r * 2 + 1)
signal <- construct_periodics(N, A, rates.true, freqs.true, phases.true, complex.signal)
```

## Parameters grid for 1d case
```{r}
if (!is.matrix(signal)) {
groups_ssa <- list(1:max.r)

@@

params_grid_ssa <- list(sd = sds, L_ssa = Ls_ssa)
params_grid_hossa <- list(sd = sds,
                          td_hossa =  trunc_dims,
                          L_hossa = tens.dims_list)
}
```

## Parameters grid for mv case
```{r}
if (is.matrix(signal)) {
groups_mssa <- list(1:max.r)

if (n_case == "") {
  Ls_mssa <- seq(from = max.r + 1,
                to = N - max.r - 1,
                by = 2)
} else {
  Ls_mssa <- seq(from = max.r + 1,
                to = N - max.r - 1,
                by = 10)
}

max.r3 <- tens_mssa_decompose(signal, N %/% 2)$Z@data |> zapsmall() |> apply(3, \(x) any(x != 0)) |> sum()
groups_homssa <- list(1:max.r)
groups3_homssa <- list(1:max.r3)

Ls_homssa = seq(from = max.r + 1,
                to = (N + 1) %/% 2,
                by = 2)

params_grid_ssa <- list(sd = sds, L_mssa = Ls_mssa)
params_grid_hossa <- list(sd = sds,
                          L_homssa = Ls_homssa)
}
```
